---
title: "Weeks 10-15"
---

# Week 10

- Review project details and get updates
- Discuss project 2
- Q&A Session

# Week 11

Tuesday:  Gain clarity on Tuesday for final third of semester

Thursday:  Project demonstration and presentation

Introduce Project 3


# Week 12

## Stan introduction via brms

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| error: FALSE
#| output: false
#| cache: false

library(brms)
library(tidyverse)

# simulate
set.seed(123)
n <- 100
x <- rnorm(n)
y <- 2*x + rnorm(n)
data <- data.frame(x, y)

# fit
fit <- brm(y ~ x, data = data, chains = 1, iter = 100)

```

```{r}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| error: FALSE

# print the fit
summary(fit)
```


```{r stcode}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| error: FALSE
#| class-output: stan

# extract stan code from brms

writeLines(stancode(fit))
```

```{r stdata}
#| echo: TRUE
#| message: FALSE
#| warning: FALSE
#| error: FALSE
#| cache: false

# extract stan data from brms
summary(standata(fit))
```


## Real life examples and applications - unique to Bayesian analyses

- [ ] [A new look at p-values for RCTs](http://www.stat.columbia.edu/~gelman/research/published/pval_RCTs_rev3.pdf)
- [ ] [A novel clinical trial](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7826635/)
- [ ] [Clinical nutrition research](https://pubmed.ncbi.nlm.nih.gov/37715007/)



# Week 13

Playing with stan
The rudiments of the stan syntax and structure

# Week 14

Real life examples

# Week 15

Final project presentation
